---
title: 'Peer-graded Assignment: Prediction Assignment Writeup'
author: "Shannah Aberin"
date: "July 31, 2019"
output: html_document
---

##Load Necessary Libraries
```{r}
library(caret)
library(rpart)
library(rpart.plot)
library(RColorBrewer)
library(rattle)
library(randomForest)
library(ggcorrplot)
```
##Load and Save data set as training and test set and explore the datasets
```{r}
train_in <- read.csv('./pml-training.csv', header=T)
valid_in <- read.csv('./pml-testing.csv', header=T)
dim(train_in)
```
##Impute the missing values to 0s and split the train set to training and validation set
```{r}
trainData<- train_in[, colSums(is.na(train_in)) == 0]
validData <- valid_in[, colSums(is.na(valid_in)) == 0]
dim(trainData)
```
##Drop unneeded variables
```{r}
trainData <- trainData[, -c(1:7)]
validData <- validData[, -c(1:7)]
dim(trainData)

```
##Prepare the training data set by partitioning it 
####70(train)/30(validation)before training the model
```{r}
set.seed(1234) 
inTrain <- createDataPartition(trainData$classe, p = 0.7, list = FALSE)
trainData <- trainData[inTrain, ]
testData <- trainData[-inTrain, ]
dim(trainData)

```
##Remove variables that have near zero variance
```{r}
zeros <- nearZeroVar(trainData)
trainData <- trainData[, -zeros]
testData  <- testData[, -zeros]
dim(trainData)
```
##Visualize the correlation of each variable using corrplot
```{r}
cor_mat <- cor(trainData[, -53])
ggcorrplot(cor_mat,  type = "upper", hc.order = TRUE, colors = brewer.pal(n = 3, name = "RdYlBu"))
```

##Model development of decision tree to for prediction using fancyRplot
```{r}
set.seed(12345)
decisionTreeMod1 <- rpart(classe ~ ., data=trainData, method="class")
fancyRpartPlot(decisionTreeMod1)
```
##Time to Predict!!!
```{r}
predictTreeMod1 <- predict(decisionTreeMod1, testData, type = "class")
```
##Display the confusion matrix of the results of the model to 
#identify the ratio of true positive and true negatives
```{r}
cmtree <- confusionMatrix(predictTreeMod1, testData$classe)
cmtree
```
## plot matrix results
```{r}
plot(cmtree$table, col = cmtree$byClass, 
     main = paste("Decision Tree - Accuracy =", round(cmtree$overall['Accuracy'], 4)))
```
##Time to predict with Random Forest Model and 
###used cross validation with 3 folds dor training the Random Forest Model
```{r}
controlRF <- trainControl(method="cv", number=3, verboseIter=FALSE)
modRF1 <- train(classe ~ ., data=trainData, method="rf", trControl=controlRF)
modRF1$finalModel
predictRF1 <- predict(modRF1, newdata=testData)
cmrf <- confusionMatrix(predictRF1, testData$classe)
cmrf
plot(modRF1)
```
#Use the best model for the test set
```{r}
Results <- predict(modRF1, newdata=validData)
Results
```

